---
title: "Daily ER visits relative to storm effects"
output: html_document
editor_options:
  chunk_output_type: console
---

## Create a storm effects model

```{r}
library(sf)
library(dplyr)

storm_intensity <- 34 # all tropical storms and hurricanes
begin_year <- 2004
end_year <- 2022

L <- "https://www.ncei.noaa.gov/data/international-best-track-archive-for-climate-stewardship-ibtracs/v04r00/access/shapefile/IBTrACS.NA.list.v04r00.lines.zip"
  
if(!"IBTrACS.NA.list.v04r00.lines.zip" %in% list.files(here::here("data"))) {
download.file(url = L,
              destfile = here::here("data",
                                    "IBTrACS.NA.list.v04r00.lines.zip"))
unzip(here::here("data", "IBTrACS.NA.list.v04r00.lines.zip"),
      exdir = here::here("data"))
}

Tracks.sf <- st_read(dsn = here::here("data"), 
                     layer = "IBTrACS.NA.list.v04r00.lines") |>
  st_transform(crs = 32616)

Tracks.sf <- Tracks.sf |>
  filter(year >= begin_year & year <= end_year) |>
  filter(USA_WIND >= storm_intensity) |>
  select(SID, SEASON, year, month, day, hour, min,
         NAME, SUBBASIN, ISO_TIME, USA_WIND, USA_PRES, USA_RMW, USA_EYE, USA_ROCI)

Tracks.sf |>
  st_drop_geometry() |>
  summarize(avgRMW = mean(USA_RMW, na.rm = TRUE) * 1.852,
                   avgEYE = mean(USA_EYE, na.rm = TRUE) * 1.852,
                   avgROCI = mean(USA_ROCI, na.rm = TRUE) * 1.852)

# Fill in missing RMW values. Start with pedigree, then use minimum pressure, and finish again with pedigree
Tracks.sf <- Tracks.sf |>
  group_by(SID) |>  # pedigree
  mutate(USA_RMW = ifelse(is.na(USA_RMW), mean(USA_RMW, na.rm = TRUE), USA_RMW))

Tracks.sf <- Tracks.sf |>
  group_by(USA_PRES) |> # minimum pressure
  mutate(USA_RMW = ifelse(is.na(USA_RMW), mean(USA_RMW, na.rm = TRUE), USA_RMW))

Tracks.sf <- Tracks.sf |>
  group_by(SID) |> # pedigree
  mutate(USA_RMW = ifelse(is.na(USA_RMW), mean(USA_RMW, na.rm = TRUE), USA_RMW))

#Add a buffer to the tracks to make segmented wind swaths
Swaths.sf <- Tracks.sf |>
  st_buffer(dist = Tracks.sf$USA_RMW * 1852) # 1852 converts to meters

# Wind swaths that cross Florida. `USAboundaries` package no longer maintained on CRAN
#devtools::install_github("ropensci/USAboundariesData")
#devtools::install_github("ropensci/USAboundaries", force = TRUE)
#install.packages("USAboundariesData", repos = "https://ropensci.r-universe.dev", type = "source")
Boundaries.sf <- USAboundaries::us_states(resolution = "low", states = "FL") |> 
  st_transform(crs = 32616)

X <- Swaths.sf |>
  st_intersects(Boundaries.sf, sparse = FALSE) #Does the swath intersect the state border?
Swaths.sf <- Swaths.sf[X, ]
Swaths.sf <- Swaths.sf |>
  mutate(Date = lubridate::as_date(ISO_TIME)) #Add a date column

# Extract the boundaries of storm impacts by storm ID. Dissolve the overlap borders of the individual swaths by storm to create a single storm swath. Add the storm category (Saffir-Simpson) based on wind speed
Swaths.sf <- Swaths.sf |>
  group_by(SID) |>
  summarize(Date0 = first(Date),
            NAME = first(NAME),
            Wind = max(USA_WIND),
            geometry = st_union(geometry)) |>
  mutate(Storm_Category = case_when(
    Wind >= 34 & Wind <= 63 ~ 0,
    Wind >= 64 & Wind <= 82 ~ 1,
    Wind >= 83 & Wind <= 95 ~ 2,
    Wind >= 96 & Wind <= 112 ~ 3,
    Wind >= 113 & Wind <= 136 ~ 4,
    Wind >= 137 ~ 5
  ))

# Transform the geometry to a geographic CRS (4326) and unionize the swaths
sf_use_s2(TRUE)

Swaths.sf <- Swaths.sf |>
    st_transform(crs = 4326)
union_Swaths2.sfg <- Swaths.sf |>
  st_union()

sum(table(Swaths.sf$Storm_Category))
table(Swaths.sf$Storm_Category)
table(lubridate::year(Swaths.sf$Date0))
```

44
 0  1  2  3  4  5 
30  3  3  4  2  2 
2004 2005 2006 2007 2008 2009 2010 2011 2012 2013 2016 2017 2018 2019 
   5    6    2    1    1    2    2    1    3    1    3    3    3    1 
2020 2021 2022 
   3    4    3

## Expand the swaths

```{r}
library(tidyr)
library(lubridate)

deltaT <- 1   # One day increment
n_new <- 8    # Number of days before & after impact

TIC.sf <- Swaths.sf |>
    rowwise() |>
    mutate(new_features = list(tidyr::tibble(
           Date = Date0 + (-n_new:n_new) * deltaT))) |>
    unnest(new_features) |>
    ungroup() |>
    select(Date, Storm_Name = NAME, Storm_Category)

TIC <- rep(rep(c("Threat", "Impact", "Cleanup"), times = c(n_new, 1, n_new)), 
           times = nrow(Swaths.sf))
TIC.sf$Storm_Effect <- TIC

TIC.sf <- TIC.sf |>
  mutate(Month = month(Date),
         Year = year(Date))
storm_months <- unique(TIC.sf$Month)
storm_year_range <- range(TIC.sf$Year)
```

## Build a zip x date effect calendar

```{r}
library(data.table)
library(tigris)
options(tigris_use_cache = TRUE)

# --- ZCTAs for Florida (2020) ---
zcta_us <- zctas(cb = TRUE, year = 2020) %>%
  select(zip = ZCTA5CE20, geometry)

fl <- states(cb = TRUE, year = 2020) %>%
  filter(STUSPS == "FL") %>%
  st_transform(st_crs(zcta_us))

zctas_fl <- st_join(zcta_us, fl, join = st_intersects, left = FALSE) %>%
  select(zip, geometry)

#Align coordinate reference systems
tic_aligned  <- st_transform(TIC.sf,  st_crs(zctas_fl))

# Spatial join: which zips intersect each TIC polygon (by date & effect)
# Keep only the columns we need from TIC
tic_keep <- tic_aligned |> 
  select(Date, Storm_Effect, geometry)

# Point-in-polygon for polygons
zcta_date_effect_sf <- st_join(
  zctas_fl,
  tic_keep,
  join = st_intersects,
  left = FALSE   # keep only zips that intersect a TIC polygon on that date
)

# Drop geometry for the calendar
zcta_date_effect <- as.data.table(st_drop_geometry(zcta_date_effect_sf))
setnames(zcta_date_effect, c("zip","Date","Storm_Effect"),
         c("zip","date","effect"))

# Normalize types
zcta_date_effect[, `:=`(
  zip    = as.character(zip),
  date   = as.IDate(date),
  effect = as.character(effect)
)]

# Resolve multiple effects on the same zip-date: Choose most severe ranking (None < Cleanup < Threat < Impact)
sev <- data.table(effect = c("None","Cleanup","Threat","Impact"),
                  rank   = 0:3)
zde <- sev[zcta_date_effect, on = "effect"]

# Keep the single most-severe label per ZIP–date
zde <- zde[ , .SD[which.max(rank)], by = .(zip, date)][ , .(zip, date, effect)]

# Build the full zip x date data frame & fill missing days with "None"
all_zips  <- sort(unique(zde$zip))
all_dates <- as.IDate(seq(as.Date("2004-12-31"), as.Date("2022-12-31"), by = "day"))
grid <- CJ(zip = as.character(all_zips), date = all_dates)

# Left-join the observed effects; fill NA → "None"
effect_calendar <- zde[grid, on = .(zip, date)]
effect_calendar[is.na(effect), effect := "None"]

# Keep effect as a factor with ordered levels
effect_calendar[, effect := factor(effect, levels = c("None","Cleanup","Threat","Impact"))]

# Exactly one effect per ZIP–day?
stopifnot(effect_calendar[, .N, by = .(zip, date)][, all(N == 1)])

# Effects present?
effect_calendar[, table(effect)]
```

effect
   None Cleanup  Threat  Impact 
6740950   64230   64240    8030 


## Merge with emergency room visits

```{r}
ER_visits.dt <- data.table::fread(here::here("data", "Combined_ER_Data.csv"))
ER_visits.df <- ER_visits.dt |>
  as.data.frame()

setDT(ER_visits.df)

# Pad ZipCode to 5 chars; drop non-sensical codes (0/NA/too short/too big)
ER_visits.df[
  , ZipCode := fifelse(ZipCode > 0 & ZipCode < 1e6, sprintf("%05d", ZipCode), NA_character_)
]
ER_visits.df[, Date := as.IDate(Date)]

ER_visits_clean <- ER_visits.df[!is.na(ZipCode) & nchar(ZipCode) == 5]

# Standardize names; optionally collapse duplicates
setnames(ER_visits_clean, c("ZipCode","Date","Total_Visits"),
                           c("zip",    "date","er_visits"))

# If there can be multiple ER rows per (zip,date), aggregate to daily total
ER_visits_clean <- ER_visits_clean[, .(er_visits = sum(er_visits, na.rm = TRUE)),
                                   by = .(zip, date)]

merged <- merge(
  effect_calendar,
  ER_visits_clean,
  by = c("zip", "date"),
  all.x = TRUE  # keep all ZIP–dates from the calendar
)

merged[is.na(er_visits), er_visits := 0L]

library(lubridate)   # for yearmonth(), wday()
setDT(merged)

# Add day of week (factor, ordered Monday–Sunday)
merged[, dow := lubridate::wday(date, label = TRUE, abbr = TRUE)]

# Add year-month (IDate stored as "YYYY-MM-01" for convenience)
merged[, year_month := as.IDate(floor_date(date, unit = "month"))]
```

## Model using logit and Poisson separately

A two-part “threshold (≤6) + truncated count (>6)” approximation (stays in `fixest`). The process is modeled as: Part A (logit): probability the true count is ≤6 (we observe 6) and part B (Poisson on truncated support): conditional mean for counts >6

```{r}
library(fixest)

dat <- merged %>%
  mutate(
    is_le6 = as.integer(er_visits == 6L)
  ) %>%
  filter(month(date) %in% 5:11)

dat_season <- as.data.table(dat)
dat_season[, row_id := .I]

mA <- feglm(
  is_le6 ~ i(effect, ref="None") | zip + year_month + dow,
  data = dat_season,
  family = binomial("logit"),
  cluster = ~ zip + year_month
)

d_gt6 <- filter(dat_season, er_visits > 6L)
mB <- fepois(
  er_visits ~ i(effect, ref="None") | zip + year_month + dow,
  data = d_gt6, cluster = ~ zip + year_month
)

summary(mA); summary(mB)
```

NOTE: 345/0/0 fixed-effects (1,328,940 observations) removed because of only 0 or only 1 outcomes or singletons.
Binomial model:
GLM estimation, family = binomial, Dep. Var.: is_le6
Observations: 2,588,544
Fixed-effects: zip: 672,  year_month: 126,  dow: 7
Standard-errors: Clustered (zip & year_month) 
                 Estimate Std. Error   z value   Pr(>|z|)    
effect::Cleanup  0.007657   0.027531  0.278111 7.8093e-01    
effect::Threat  -0.004921   0.030158 -0.163164 8.7039e-01    
effect::Impact   0.194508   0.047288  4.113296 3.9005e-05 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Log-Likelihood: -406,796.1   Adj. Pseudo R2: 0.150202
           BIC:  825,494.1     Squared Cor.: 0.052679

Interpretation
Cleanup and Threat days: No meaningful evidence that these periods change the probability of low ER visits.
Impact days (landfall): A positive and significant log-odds means ER visit counts are more likely to fall in the ≤6 range. In other words, ER utilization tends to drop during direct storm impact

This makes sense: Many people stay home. Clinics/hospitals reduce service hours. Travel becomes difficult.
This model focuses on the lower tail of demand and here, Impact days show a clear reduction

NOTE: 8/0/0 fixed-effects (8 observations) removed because of only 0 outcomes or singletons.
Poisson model
Poisson estimation, Dep. Var.: er_visits
Observations: 2,722,224
Fixed-effects: zip: 912,  year_month: 126,  dow: 7
Standard-errors: Clustered (zip & year_month) 
                 Estimate Std. Error  z value  Pr(>|z|)    
effect::Cleanup  0.023700   0.008545  2.77347 0.0055461 ** 
effect::Threat   0.018231   0.008404  2.16917 0.0300695 *  
effect::Impact  -0.085411   0.026034 -3.28079 0.0010352 ** 
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Log-Likelihood: -8,500,999.7   Adj. Pseudo R2: 0.604913
           BIC: 17,017,498.0     Squared Cor.: 0.880416


Interpretation
Cleanup days: Shows a small but significant +2–3% ER increase. This makes intuitive sense as delayed care resumes injuries from debris removal reconnecting the healthcare system after outages. Threat days (pre-impact conditions): ER visits increase (+1–2%). People may seek care before the storm hits (filling prescriptions, addressing minor issues early)
Impact days: ER visits drop sharply (≈ –8%). This lines up perfectly with the binomial model’s finding: ER demand suppresses during the storm’s immediate impact

Jointly the models show storm Impact periods suppress ER utilization — likely due to mobility constraints, closures, and shelter-in-place behavior. Pre-storm Threat periods elevate ER visits, consistent with precautionary care. Cleanup periods also show elevated ER visits as expected with resuming delayed care, debris removal injuries and reconnecting with heathcare system after outages. This is exactly the kind of pattern you’d expect in a population with both anticipatory behavior and acute disruption around major storms

During hurricane season in Florida, patients ramp up ER usage slightly the week before storms (Threat). On the direct storm impact day, ER demand collapses (Impact). After storms, ER visits rebound and exceed normal levels (Cleanup). This is a classic displacement–disruption pattern. Pre-storm increase → mid-storm suppression → post-storm rebound. The fixed effects ensure this is not confounded by ZIP characteristics, seasonal cycles, or weekly patterns

## Compare results with a subset of the stronger storms 55+

GLM estimation, family = binomial, Dep. Var.: is_le6
Observations: 2,576,988
Fixed-effects: zip: 669,  year_month: 126,  dow: 7
Standard-errors: Clustered (zip & year_month) 
                 Estimate Std. Error   z value Pr(>|z|)    
effect::Cleanup -0.111066   0.075692 -1.467345 0.142282    
effect::Threat  -0.032148   0.056498 -0.569022 0.569341    
effect::Impact   0.261600   0.136040  1.922963 0.054485 .  
Log-Likelihood: -392,643.5   Adj. Pseudo R2: 0.153986
           BIC:  797,140.9     Squared Cor.: 0.053344
           
Poisson estimation, Dep. Var.: er_visits
Observations: 2,648,475
Fixed-effects: zip: 895,  year_month: 126,  dow: 7
Standard-errors: Clustered (zip & year_month) 
                 Estimate Std. Error   z value  Pr(>|z|)    
effect::Cleanup  0.061417   0.023028  2.667050 0.0076520 ** 
effect::Threat   0.014805   0.019471  0.760378 0.4470288    
effect::Impact  -0.235362   0.077399 -3.040881 0.0023589 ** 
Log-Likelihood: -8,280,629.7   Adj. Pseudo R2: 0.605412
           BIC: 16,576,477.7     Squared Cor.: 0.880557
           
## Unified expected ER visit predictions

```{r}
# -------------------------------------------------------------------
# 1. Predict P(Y <= 6) from the logit part
# -------------------------------------------------------------------
dat_season[, prob_le6 := predict(
  mA,
  type    = "response",
  newdata = dat_season
)]

# -------------------------------------------------------------------
# 2. Predict E[Y | Y > 6] from the Poisson part, evaluated on all rows
# -------------------------------------------------------------------
dat_season[, mu_gt6 := predict(
  mB,
  type    = "response",
  newdata = dat_season
)]

# -------------------------------------------------------------------
# 3. Expected value for <=6 cases  
# Since <=6 are top-coded as 6, we treat E[Y | Y <= 6] = 6
# -------------------------------------------------------------------
mean_le6 <- 6   # or mean(dat_season[er_visits <= 6]$er_visits) if ever unmasked

# -------------------------------------------------------------------
# 4. Combined expected ER visits for each observation
# E[Y] = P(Y<=6)*E[Y<=6] + (1-P(Y<=6))*E[Y|Y>6]
# -------------------------------------------------------------------
dat_season[, E_Y := prob_le6 * mean_le6 +
                      (1 - prob_le6) * mu_gt6]

# -------------------------------------------------------------------
# 5. Average expected visits by effect (Cleanup, Threat, Impact, None)
# -------------------------------------------------------------------
combined_effect <- dat_season[
  , .(expected_visits = mean(E_Y, na.rm = TRUE)),
  by = effect
][order(effect)]

combined_effect
```

    effect expected_visits
    <fctr>           <num>
1:    None        13.63772
2: Cleanup        14.20271
3:  Threat        14.27728
4:  Impact        12.92706

## Average only over ever‑exposed ZIPs

Storm‑exposed ZIPs” = ZIPs that had any non‑None effect at least once during May–Nov across the study period

```{r}
# --- Assumes you already have:
# dat_season (data.table): zip, date, er_visits, effect, year_month, dow, etc.
# mA (logit) and mB (Poisson) fitted on dat_season
# factor levels for effect: c("None","Cleanup","Threat","Impact")

# 1) Identify ever-exposed ZIPs from the model sample
exposed_zips <- unique(dat_season[effect != "None", zip])

# 2) Predict on the estimation sample (SAFE: newdata = dat_season)
dat_season[, prob_le6 := predict(mA, type = "response", newdata = dat_season)]
dat_season[, mu_gt6   := predict(mB, type = "response", newdata = dat_season)]

# 3) Combine to a single expected outcome per row
mean_le6 <- 6   # If ≤6 are truly top-coded as 6. Replace with empirical mean if available.
dat_season[, E_Y := prob_le6 * mean_le6 + (1 - prob_le6) * mu_gt6]

# 4) Restrict averaging SET to ever-exposed ZIPs only
dat_exposed <- dat_season[zip %in% exposed_zips]

# 5) Descriptive average expected visits by observed effect (Cleanup/Threat/Impact/None)
combined_effect_exposed <- dat_exposed[
  , .(expected_visits = mean(E_Y, na.rm = TRUE)),
  by = effect
][order(effect)]

combined_effect_exposed
```

What would the average ER visits have been among storm‑exposed ZIPs if we set every row’s effect to Threat/Impact/Cleanup/None?

```{r}
# Helper: counterfactual mean restricting the averaging SET to exposed ZIPs
counterfactual_mean_exposed <- function(level){
  newdat <- data.table::copy(dat_season)
  newdat[, effect := factor(level, levels = c("None","Cleanup","Threat","Impact"))]
  newdat[, prob_le6 := predict(mA, type = "response", newdata = newdat)]
  newdat[, mu_gt6   := predict(mB, type = "response", newdata = newdat)]
  newdat[, EY := prob_le6 * mean_le6 + (1 - prob_le6) * mu_gt6]
  # Average only over the ever-exposed ZIPs
  mean(newdat[zip %in% exposed_zips, EY], na.rm = TRUE)
}

levels_eff <- c("None","Cleanup","Threat","Impact")
cf_means_exposed <- sapply(levels_eff, counterfactual_mean_exposed)
diff_vs_none_exposed <- cf_means_exposed - cf_means_exposed["None"]

cf_means_exposed
diff_vs_none_exposed
```

## Average only over “exposed‑today” ZIP‑days

Here, the averaging set is only those ZIP‑days with effect != "None" in the observed data (i.e., actually in a Threat/Impact/Cleanup window). This will yield larger magnitudes, because it focuses on the storm windows

```{r}
# Identify the exposed-today subset in the observed data
dat_exposed_today <- dat_season[effect != "None"]

# Descriptive mean by observed effect (on exposed days only)
combined_effect_exposed_today <- dat_exposed_today[
  , .(expected_visits = mean(E_Y, na.rm = TRUE)),
  by = effect
][order(effect)]

combined_effect_exposed_today
```

Counterfactual within “exposed‑today”: Average E[Y(a)] only over rows that were exposed today in the observed data

```{r}
counterfactual_mean_exposed_today <- function(level){
  newdat <- data.table::copy(dat_season)
  newdat[, effect := factor(level, levels = c("None","Cleanup","Threat","Impact"))]
  newdat[, prob_le6 := predict(mA, type = "response", newdata = newdat)]
  newdat[, mu_gt6   := predict(mB, type = "response", newdata = newdat)]
  newdat[, EY := prob_le6 * mean_le6 + (1 - prob_le6) * mu_gt6]
  # Average only over the rows that were non-None originally
  mean(newdat[dat_season$effect != "None", EY], na.rm = TRUE)
}

cf_means_exposed_today <- sapply(levels_eff, counterfactual_mean_exposed_today)
diff_vs_none_exposed_today <- cf_means_exposed_today - cf_means_exposed_today["None"]

cf_means_exposed_today
diff_vs_none_exposed_today
```

## Function: two‑part g‑computation with clustered bootstrap + figure

```{r}
library(data.table)
library(dplyr)
library(lubridate)
library(fixest)
library(ggplot2)

two_part_gcomp_bootstrap <- function(
  df,
  effect_levels = c("None", "Cleanup", "Threat", "Impact"),
  season_months = 5:11,
  mean_le6      = 6,
  ref_level     = "None",
  averaging_set = c("all","ever_exposed","exposed_today"),
  n_boot        = 200L,
  cluster       = c("zip","zip_year_month"),
  seed          = 123,
  verbose       = TRUE
) {

  averaging_set <- match.arg(averaging_set)
  cluster       <- match.arg(cluster)

  # ---- 0) Data prep ----
  dat <- as.data.table(df)
  dat[, date := as.IDate(date)]
  dat[, zip  := as.character(zip)]
  dat[, dow        := lubridate::wday(date, label = TRUE, abbr = TRUE)]
  dat[, year_month := as.IDate(floor_date(date, "month"))]
  dat[, effect     := factor(effect, levels = effect_levels)]
  dat <- dat[month(date) %in% season_months]
  dat[, is_le6 := as.integer(er_visits == 6L)]
  dat[, row_id := .I]

  ever_exposed_zips <- unique(dat[effect != "None", zip])
  exposed_today_id  <- dat[, effect != "None"]

  # ---- 1) Fit two-part models (SUPPRESS fixest NOTE/WARNING) ----
  fA <- as.formula(paste0("is_le6 ~ i(effect, ref='", ref_level, "') | zip + year_month + dow"))
  fB <- as.formula(paste0("er_visits ~ i(effect, ref='", ref_level, "') | zip + year_month + dow"))

  mA <- suppressMessages(suppressWarnings(
    feglm(fA, data = dat, family = binomial("logit"),
          cluster = ~ zip + year_month)
  ))

  d_gt6 <- dat[er_visits > 6L]

  mB <- suppressMessages(suppressWarnings(
    fepois(fB, data = d_gt6, cluster = ~ zip + year_month)
  ))

  # ---- averaging set helper ----
  get_avg_mask <- function(dat0) {
    if (averaging_set == "all")           return(rep(TRUE, nrow(dat0)))
    if (averaging_set == "ever_exposed")  return(dat0$zip %in% ever_exposed_zips)
    if (averaging_set == "exposed_today") return(dat0$effect != "None")
  }

  avg_mask <- get_avg_mask(dat)

  # ---- 2) Counterfactual mean ----
  counterfactual_mean <- function(level, modA, modB) {
    newdat <- copy(dat)
    newdat[, effect := factor(level, levels = effect_levels)]

    newdat[, prob_le6 := predict(modA, type = "response", newdata = newdat)]
    newdat[, mu_gt6   := predict(modB, type = "response", newdata = newdat)]
    newdat[, EY := prob_le6 * mean_le6 + (1 - prob_le6) * mu_gt6]

    mean(newdat[avg_mask, EY], na.rm = TRUE)
  }

  set.seed(seed)
  cf_levels <- effect_levels
  cf_point <- sapply(cf_levels, counterfactual_mean, modA = mA, modB = mB)
  names(cf_point) <- cf_levels
  diff_point <- cf_point - cf_point[ref_level]

  # ---- 3) Setup bootstrap clusters ----
  if (cluster == "zip") {
    dat[, boot_cluster := zip]
  } else {
    dat[, boot_cluster := interaction(zip, year_month, drop = TRUE)]
  }

  clust_vals <- unique(dat$boot_cluster)
  n_clusters <- length(clust_vals)

  # ---- 4) Bootstrap replicate function WITH PROGRESS COUNTER ----
  do_one_boot <- function(b) {

    if (verbose && b %% 10 == 0)
      cat("Bootstrap replicate:", b, "of", n_boot, "\n")

    set.seed(seed + b)
    draw <- sample(clust_vals, n_clusters, replace = TRUE)

    dat_b <- rbindlist(lapply(draw, function(z) dat[boot_cluster == z]), use.names = TRUE)
    dat_b[, effect := factor(effect, levels = effect_levels)]

    d_gt6_b <- dat_b[er_visits > 6L]
    if (nrow(d_gt6_b) == 0) return(rep(NA_real_, length(cf_levels)))

    mA_b <- suppressMessages(suppressWarnings(
      feglm(fA, data = dat_b, family = binomial("logit"),
            cluster = ~ zip + year_month)
    ))

    mB_b <- suppressMessages(suppressWarnings(
      fepois(fB, data = d_gt6_b, cluster = ~ zip + year_month)
    ))

    if (is.null(mA_b) || is.null(mB_b)) return(rep(NA_real_, length(cf_levels)))

    # averaging mask for the bootstrap sample
    avg_mask_b <- get_avg_mask(dat_b)

    newdat <- copy(dat_b)
    sapply(cf_levels, function(level) {
      newdat[, effect := factor(level, levels = effect_levels)]
      newdat[, prob_le6 := predict(mA_b, type = "response", newdata = newdat)]
      newdat[, mu_gt6   := predict(mB_b, type = "response", newdata = newdat)]
      newdat[, EY := prob_le6 * mean_le6 + (1 - prob_le6) * mu_gt6]
      mean(newdat[avg_mask_b, EY], na.rm = TRUE)
    })
  }

  # ---- run bootstrap ----
  boot_mat <- t(replicate(n_boot, do_one_boot(sample.int(1e9,1))))
  colnames(boot_mat) <- cf_levels
  boot_df <- as.data.table(boot_mat)
  boot_df <- boot_df[rowSums(is.na(boot_df)) < length(cf_levels)]

  # Differences vs reference
  boot_diff <- boot_df[, lapply(.SD, function(v) v - boot_df[[ref_level]])]

  # ---- 5) CI computation ----
  pct <- function(x,p) quantile(x, probs = p, na.rm = TRUE)

  ci_tbl_means <- data.table(
    effect   = cf_levels,
    estimate = cf_point,
    ci_lo    = sapply(cf_levels, function(k) pct(boot_df[[k]], 0.025)),
    ci_hi    = sapply(cf_levels, function(k) pct(boot_df[[k]], 0.975))
  )

  diff_levels <- setdiff(cf_levels, ref_level)
  ci_tbl_diffs <- data.table(
    effect   = diff_levels,
    estimate = diff_point[diff_levels],
    ci_lo    = sapply(diff_levels, function(k) pct(boot_diff[[k]], 0.025)),
    ci_hi    = sapply(diff_levels, function(k) pct(boot_diff[[k]], 0.975))
  )

  # ---- 6) Publication-ready plot ----
  plot_df <- ci_tbl_diffs %>%
    mutate(effect = factor(effect, levels = c("Threat","Impact","Cleanup"))) %>%
    arrange(effect)

  fig <- ggplot(plot_df, aes(x = effect, y = estimate)) +
    geom_hline(yintercept = 0, color = "grey50") +
    geom_pointrange(aes(ymin = ci_lo, ymax = ci_hi),
                    linewidth = 0.6, fatten = 2, color = "#2C3E50") +
    labs(
      x = NULL,
      y = "Difference in expected ER visits vs None",
      title    = "Combined Storm-Phase Effect on Expected ER Visits",
      subtitle = paste(
        "Averaging set:", averaging_set,
        "|", n_boot, "bootstrap replicates",
        "| clustered by", cluster
      ),
      caption  = "Two-part model (logit + Poisson) combined via g-computation"
    ) +
    theme_minimal(base_size = 12) +
    theme(plot.title.position="plot",
          panel.grid.minor = element_blank())

  list(
    models = list(mA = mA, mB = mB),
    point_estimates = list(
      expected = ci_tbl_means,
      diff_vs_none = ci_tbl_diffs
    ),
    bootstrap = list(
      cf = boot_df,
      diff = boot_diff
    ),
    figure = fig
  )
}
```

## Combine models and create figure

```{r}
# Build the modeling dataset by joining:
dat_model <- merge(
  effect_calendar[, .(zip, date, effect)],
  ER_visits_clean[, .(zip, date, er_visits)],
  by = c("zip","date"),
  all.x = TRUE
)
dat_model[is.na(er_visits), er_visits := 0L]

# Run (ZIP-clustered bootstrap)
res <- two_part_gcomp_bootstrap(
  df            = dat_model,
  averaging_set = "exposed_today", # all, exposed_today"
  n_boot        = 200,
  cluster       = "zip"
)

# View the unified, combined effect (expected ER visits) with CIs
res$point_estimates$expected
res$point_estimates$diff_vs_none

# View the key plot (differences vs "None")
print(res$figure)
```

    effect estimate    ci_lo    ci_hi
    <char>    <num>    <num>    <num>
1:    None 13.96969 13.43418 14.46346
2: Cleanup 14.25720 13.73704 14.77212
3:  Threat 14.23676 13.69721 14.75458
4:  Impact 12.82809 12.29979 13.31213

    effect  estimate      ci_lo      ci_hi
    <char>     <num>      <num>      <num>
1: Cleanup  0.287507  0.2357466  0.3401862
2:  Threat  0.267074  0.2185509  0.3241192
3:  Impact -1.141603 -1.2531281 -1.0193433

```{r}
# Save outputs
data.table::fwrite(res$point_estimates$expected,
                   here::here("output","expected_visits_by_effect.csv"))
data.table::fwrite(res$point_estimates$diff_vs_none,
                   here::here("output","diff_vs_none_by_effect.csv"))
ggplot2::ggsave(filename = here::here("figs","combined_effect_diffs_vs_none.png"),
                plot = res$figure, width = 7, height = 4.5, dpi = 300)
```

## Get zip population table

```{r}
# --- Packages & key ---
library(tidycensus)
library(dplyr)
library(janitor)
library(purrr)

# Make sure your key is available (run once with install=TRUE; then restart R)
# census_api_key("YOUR_API_KEY_HERE", install = TRUE)
stopifnot(Sys.getenv("CENSUS_API_KEY") != "")

# --- One-year example (2022) ---
df_2022 <- get_acs(
  geography   = "zcta",           # don't pass state= for ZCTAs
  variables   = "B01003_001",     # total population
  year        = 2022,
  survey      = "acs5",
  geometry    = FALSE,
  cache_table = TRUE
) %>%
  clean_names() %>%
  transmute(
    zcta = geoid,
    pop  = estimate,
    year = 2022
  )

head(df_2022)

acs_years <- 2009:2024  # ACS 5-year is available starting 2009

get_zcta_pop_year <- function(yr) {
  # small pause to avoid throttling bursts
  Sys.sleep(0.2)
  get_acs(
    geography   = "zcta",
    variables   = "B01003_001",
    year        = yr,
    survey      = "acs5",
    geometry    = FALSE,
    cache_table = TRUE
  ) %>%
    clean_names() %>%
    transmute(
      zcta = geoid,
      pop  = estimate,
      year = yr
    )
}

safe_get <- purrr::possibly(get_zcta_pop_year,
                            otherwise = tibble(zcta = character(), pop = double(), year = integer()))

zcta_annual <- map_dfr(acs_years, safe_get)

stopifnot(all(c("zcta","pop","year") %in% names(zcta_annual)))
zcta_annual %>% count(year) %>% arrange(year)
```

Clip to Florida
```{r}
library(tigris)
library(sf)
library(data.table)
library(lubridate)
options(tigris_use_cache = TRUE)

zcta_fl <- zctas(cb = TRUE, year = 2020) %>%
  st_drop_geometry() %>%
  select(zcta = ZCTA5CE20)
fl_state <- states(cb = TRUE, year = 2020) %>% filter(STUSPS == "FL")
# spatially keep only those intersecting FL
zcta_fl <- st_join(zctas(cb = TRUE, year = 2020), fl_state, join = st_intersects, left = FALSE) %>%
  st_drop_geometry() %>%
  transmute(zcta = ZCTA5CE20)

# Keep Florida ZCTAs, build full 2004–2022 annual and monthly series
full_years <- 2004:2022

zip_annual <- zcta_annual %>%
  inner_join(zcta_fl, by = "zcta") %>%
  transmute(zip = zcta, year, pop)

zip_year_grid <- tidyr::expand_grid(zip = unique(zip_annual$zip), year = full_years)

zip_annual_full <- zip_year_grid %>%
  left_join(zip_annual, by = c("zip","year")) %>%
  arrange(zip, year) %>%
  group_by(zip) %>%
  mutate(pop = approx(x = year[!is.na(pop)], y = pop[!is.na(pop)], xout = year, 
                      method = "linear", rule = 2)$y) %>%
  ungroup()

zip_monthly <- zip_annual_full %>%
  mutate(year_month = as.IDate(ymd(paste0(year, "-01-01")))) %>%
  group_by(zip) %>%
  tidyr::uncount(12, .remove = FALSE) %>%            # repeat each year 12 times
  group_by(zip, year) %>%
  mutate(year_month = as.IDate(year_month + months(row_number()-1))) %>%
  ungroup() %>%
  select(zip, year_month, pop)

# (Optional) Limit to May–Nov
zip_monthly_season <- zip_monthly %>%
  filter(month(year_month) %in% 5:11)
```

Population offsets and weights

```{r}
# ---------------------------------------------------------------
# Two-part g-computation helper with population offsets & weights
# ---------------------------------------------------------------
library(data.table)
library(dplyr)
library(lubridate)
library(fixest)
library(ggplot2)

two_part_gcomp_bootstrap <- function(
  df,                      # panel: zip, date, er_visits, effect
  pop_dt         = NULL,   # monthly population: zip, year_month OR date, pop
  pop_by         = c("year_month","date"),
  use_poisson_offset = TRUE,           # offset = log(pop) in Poisson
  gcomp_weighting    = c("none","population"), # population-weighted averaging
  effect_levels  = c("None", "Cleanup", "Threat", "Impact"),
  season_months  = 5:11,
  mean_le6       = 6,
  ref_level      = "None",
  averaging_set  = c("all","ever_exposed","exposed_today"),
  n_boot         = 200L,
  cluster        = c("zip","zip_year_month"),
  seed           = 123,
  verbose        = TRUE
) {
  pop_by <- match.arg(pop_by)
  gcomp_weighting <- match.arg(gcomp_weighting)
  averaging_set <- match.arg(averaging_set)
  cluster       <- match.arg(cluster)

  # -- Data prep
  dat <- as.data.table(df)
  dat[, date := as.IDate(date)]
  dat[, zip  := sprintf("%05s", as.character(zip))]
  dat[, dow        := wday(date, label = TRUE, abbr = TRUE)]
  dat[, year_month := as.IDate(floor_date(date, "month"))]
  dat[, effect     := factor(effect, levels = effect_levels)]
  dat <- dat[month(date) %in% season_months]
  dat[, is_le6 := as.integer(er_visits == 6L)]
  dat[, row_id := .I]

  # -- Merge population if provided
  if (!is.null(pop_dt)) {
    pop <- as.data.table(pop_dt)
    pop[, zip := sprintf("%05s", as.character(zip))]
    if (pop_by == "year_month") {
      stopifnot("year_month" %in% names(pop))
      pop[, year_month := as.IDate(year_month)]
      dat <- pop[dat, on = c("zip","year_month")]
    } else {
      stopifnot("date" %in% names(pop))
      pop[, date := as.IDate(date)]
      dat <- pop[dat, on = c("zip","date")]
    }
    # Impute missing pop conservatively: within-zip median, else global median
    if (anyNA(dat$pop)) {
      dat[, pop := ifelse(is.na(pop), median(pop, na.rm = TRUE), pop), by = zip]
      if (anyNA(dat$pop)) dat[is.na(pop), pop := median(pop, na.rm = TRUE)]
    }
    # Guard: non-positive -> small positive to avoid -Inf offset
    dat[!is.finite(pop) | pop <= 0, pop := 1]
    dat[, log_pop := log(pop)]
  } else {
    dat[, `:=`(pop = NA_real_, log_pop = NA_real_)]
  }

  # -- Exposure sets
  ever_exposed_zips <- unique(dat[effect != "None", zip])
  exposed_today_id  <- dat[, effect != "None"]

  # -- Two-part models (suppress fixest messages/notes)
  fA <- as.formula(paste0("is_le6 ~ i(effect, ref='", ref_level, "') | zip + year_month + dow"))
  fB <- as.formula(paste0("er_visits ~ i(effect, ref='", ref_level, "') | zip + year_month + dow"))

  mA <- suppressMessages(suppressWarnings(
    feglm(fA, data = dat, family = binomial("logit"), cluster = ~ zip + year_month)
  ))

  d_gt6 <- dat[er_visits > 6L]
  if (use_poisson_offset && !is.null(pop_dt)) {
    mB <- suppressMessages(suppressWarnings(
      fepois(fB, data = d_gt6, cluster = ~ zip + year_month, offset = ~ log_pop)
    ))
  } else {
    mB <- suppressMessages(suppressWarnings(
      fepois(fB, data = d_gt6, cluster = ~ zip + year_month)
    ))
  }

  # -- Averaging mask & weights
  get_avg_mask <- function(d0) {
    if (averaging_set == "all")           return(rep(TRUE, nrow(d0)))
    if (averaging_set == "ever_exposed")  return(d0$zip %in% ever_exposed_zips)
    if (averaging_set == "exposed_today") return(d0$effect != "None")
  }
  avg_mask <- get_avg_mask(dat)

  get_weights <- function(d0, mask) {
    if (gcomp_weighting == "population" && !is.null(pop_dt)) {
      w <- d0$pop; w[!mask] <- NA_real_; return(w)
    } else {
      return(rep(1, nrow(d0)))
    }
  }
  base_weights <- get_weights(dat, avg_mask)

  wmean <- function(x, w) {
    if (all(is.na(w))) return(mean(x, na.rm = TRUE))
    sum(x * w, na.rm = TRUE) / sum(w[!is.na(x)], na.rm = TRUE)
  }

  # -- Counterfactual mean for a given effect level
  counterfactual_mean <- function(level, modA, modB, d0, mask, w) {
    nd <- copy(d0)
    nd[, effect := factor(level, levels = effect_levels)]
    nd[, prob_le6 := predict(modA, type = "response", newdata = nd)]
    nd[, mu_gt6   := predict(modB, type = "response", newdata = nd)]
    nd[, EY := prob_le6 * mean_le6 + (1 - prob_le6) * mu_gt6]
    wmean(nd$EY[mask], w[mask])
  }

  set.seed(seed)
  cf_levels <- effect_levels
  cf_point  <- sapply(cf_levels, counterfactual_mean,
                      modA = mA, modB = mB, d0 = dat,
                      mask = avg_mask, w = base_weights)
  names(cf_point) <- cf_levels
  diff_point <- cf_point - cf_point[ref_level]

  # -- Clusters for bootstrap
  if (cluster == "zip") {
    dat[, boot_cluster := zip]
  } else {
    dat[, boot_cluster := interaction(zip, year_month, drop = TRUE)]
  }
  clust_vals <- unique(dat$boot_cluster)
  n_clusters <- length(clust_vals)

  # -- Bootstrap replicate with progress every 10th
  do_one_boot <- function(b) {
    if (verbose && b %% 10 == 0) cat("Bootstrap replicate:", b, "of", n_boot, "\n")
    set.seed(seed + b)
    draw <- sample(clust_vals, n_clusters, replace = TRUE)
    db <- rbindlist(lapply(draw, function(z) dat[boot_cluster == z]), use.names = TRUE)
    db[, effect := factor(effect, levels = effect_levels)]
    d_gt6_b <- db[er_visits > 6L]
    if (nrow(d_gt6_b) == 0) return(rep(NA_real_, length(cf_levels)))

    mA_b <- suppressMessages(suppressWarnings(
      feglm(fA, data = db, family = binomial("logit"), cluster = ~ zip + year_month)
    ))
    mB_b <- if (use_poisson_offset && !is.null(pop_dt)) {
      suppressMessages(suppressWarnings(
        fepois(fB, data = d_gt6_b, cluster = ~ zip + year_month, offset = ~ log_pop)
      ))
    } else {
      suppressMessages(suppressWarnings(
        fepois(fB, data = d_gt6_b, cluster = ~ zip + year_month)
      ))
    }
    if (is.null(mA_b) || is.null(mB_b)) return(rep(NA_real_, length(cf_levels)))

    avg_mask_b <- get_avg_mask(db)
    w_b        <- get_weights(db, avg_mask_b)

    nd <- copy(db)
    sapply(cf_levels, function(level) {
      nd[, effect := factor(level, levels = effect_levels)]
      nd[, prob_le6 := predict(mA_b, type = "response", newdata = nd)]
      nd[, mu_gt6   := predict(mB_b, type = "response", newdata = nd)]
      nd[, EY := prob_le6 * mean_le6 + (1 - prob_le6) * mu_gt6]
      wmean(nd$EY[avg_mask_b], w_b[avg_mask_b])
    })
  }

  boot_mat <- t(replicate(n_boot, do_one_boot(sample.int(1e9,1))))
  colnames(boot_mat) <- cf_levels
  boot_df <- as.data.table(boot_mat)
  boot_df <- boot_df[rowSums(is.na(boot_df)) < length(cf_levels)]
  boot_diff <- boot_df[, lapply(.SD, function(v) v - boot_df[[ref_level]])]

  # -- 95% percentile CIs
  pct <- function(x,p) quantile(x, probs = p, na.rm = TRUE)
  ci_tbl_means <- data.table(
    effect   = cf_levels,
    estimate = cf_point,
    ci_lo    = sapply(cf_levels, function(k) pct(boot_df[[k]], 0.025)),
    ci_hi    = sapply(cf_levels, function(k) pct(boot_df[[k]], 0.975))
  )
  diff_levels <- setdiff(cf_levels, ref_level)
  ci_tbl_diffs <- data.table(
    effect   = diff_levels,
    estimate = diff_point[diff_levels],
    ci_lo    = sapply(diff_levels, function(k) pct(boot_diff[[k]], 0.025)),
    ci_hi    = sapply(diff_levels, function(k) pct(boot_diff[[k]], 0.975))
  )

  # -- Publication-ready plot (Threat → Impact → Cleanup)
  plot_df <- ci_tbl_diffs %>%
    mutate(effect = factor(effect, levels = c("Threat","Impact","Cleanup"))) %>%
    arrange(effect)

  ylab <- if (use_poisson_offset) {
    "Δ expected ER visits per ZIP-day (population-adjusted) vs None"
  } else if (gcomp_weighting == "population") {
    "Δ expected ER visits (population-weighted) vs None"
  } else {
    "Δ expected ER visits vs None"
  }

  fig <- ggplot(plot_df, aes(x = effect, y = estimate)) +
    geom_hline(yintercept = 0, color = "grey50") +
    geom_pointrange(aes(ymin = ci_lo, ymax = ci_hi),
                    linewidth = 0.6, fatten = 2, color = "#2C3E50") +
    labs(
      x = NULL, y = ylab,
      title    = "Combined Storm-Phase Effect on Expected ER Visits",
      subtitle = paste(
        "Averaging set:", averaging_set,
        "|", n_boot, "bootstrap replicates",
        "| clustered by", cluster,
        if (!is.null(pop_dt)) paste0("| population by ", pop_by) else ""
      ),
      caption  = "Two-part model (logit + Poisson) combined via g-computation"
    ) +
    theme_minimal(base_size = 12) +
    theme(plot.title.position="plot",
          panel.grid.minor = element_blank())

  list(
    models = list(mA = mA, mB = mB),
    point_estimates = list(expected = ci_tbl_means, diff_vs_none = ci_tbl_diffs),
    bootstrap = list(cf = boot_df, diff = boot_diff),
    figure = fig
  )
}
```

Run it

```{r}
library(dplyr)
library(tidyr)

# dat_model: merge of effect_calendar (zip, date, effect) + ER_visits_clean (zip, date, er_visits)
dat_model <- effect_calendar[, .(zip, date, effect)] %>%
  dplyr::left_join(ER_visits_clean, by = c("zip","date")) %>%
  mutate(er_visits = replace_na(er_visits, 0L))

# zip_monthly: (zip, year_month, pop) — what you just built
# Make sure types and keys are as expected
zip_monthly <- zip_monthly %>%
  mutate(zip = sprintf("%05s", zip),
         year_month = as.IDate(year_month))

# Run the two-part g-comp with population offset + population-weighted averaging
res <- two_part_gcomp_bootstrap(
  df                 = dat_model,
  pop_dt             = zip_monthly,     # <— monthly populations
  pop_by             = "year_month",    # matches the key in pop_dt
  use_poisson_offset = TRUE,            # model rates via offset = log(pop)
  gcomp_weighting    = "population",    # population-weighted g-comp averaging
  averaging_set      = "ever_exposed",  # or "exposed_today" / "all"
  n_boot             = 200,             # 500–1000 for publication
  cluster            = "zip",
  seed               = 123,
  verbose            = TRUE
)

# View results
print(res$figure)
res$point_estimates$expected
res$point_estimates$diff_vs_none
```

# Save outputs

```{r}
data.table::fwrite(res$point_estimates$expected,    here::here("output","expected_by_effect_pop.csv"))
data.table::fwrite(res$point_estimates$diff_vs_none,here::here("output","diff_vs_none_pop.csv"))
ggplot2::ggsave(here::here("figs","combined_effects_population.png"),
                res$figure, width = 7, height = 4.5, dpi = 300)
```

# Build the 3‑panel figure (Threat → Impact → Cleanup), stratified by intensity

```{r}
# --- Packages ---
library(data.table)
library(dplyr)
library(ggplot2)
library(lubridate)
library(sf)
library(tidyr)

# -------------------------------
# 1) Utility: build effect_calendar at a given wind threshold
# -------------------------------
build_effect_calendar_for_threshold <- function(Swaths_sf_full, thr_knots = 34L) {
  # Filter storms by intensity
  Swaths_t <- Swaths_sf_full %>%
    dplyr::filter(Wind >= thr_knots)

  # Rebuild the ±8 day TIC table
  deltaT <- 1L
  n_new  <- 8L

  TIC.sf <- Swaths_t %>%
    rowwise() %>%
    mutate(new_features = list(tidyr::tibble(
      Date = Date0 + (-n_new:n_new) * deltaT
    ))) %>%
    unnest(new_features) %>%
    ungroup() %>%
    select(Date, Storm_Name = NAME, Storm_Category)

  TIC <- rep(rep(c("Threat", "Impact", "Cleanup"), times = c(n_new, 1, n_new)), 
             times = nrow(Swaths_t))
  TIC.sf$Storm_Effect <- TIC

  # Spatial join (keep your original CRS joins)
  zcta_us <- tigris::zctas(cb = TRUE, year = 2020) %>% select(zip = ZCTA5CE20, geometry)
  fl      <- tigris::states(cb = TRUE, year = 2020) %>% filter(STUSPS == "FL") %>%
               st_transform(st_crs(zcta_us))

  zctas_fl <- st_join(zcta_us, fl, join = st_intersects, left = FALSE) %>%
    select(zip, geometry)

  tic_aligned <- st_transform(TIC.sf, st_crs(zctas_fl))
  tic_keep    <- tic_aligned %>% select(Date, Storm_Effect, geometry)

  zcta_date_effect_sf <- st_join(
    zctas_fl,
    tic_keep,
    join = st_intersects,
    left = FALSE
  )

  zcta_date_effect <- as.data.table(st_drop_geometry(zcta_date_effect_sf))
  setnames(zcta_date_effect, c("zip","Date","Storm_Effect"), c("zip","date","effect"))
  zcta_date_effect[, `:=`(
    zip    = as.character(zip),
    date   = as.IDate(date),
    effect = as.character(effect)
  )]

  # Resolve multiple effects per ZIP-day by severity
  sev <- data.table(effect = c("None","Cleanup","Threat","Impact"), rank = 0:3)
  zde <- sev[zcta_date_effect, on = "effect"]
  zde <- zde[, .SD[which.max(rank)], by = .(zip, date)][, .(zip, date, effect)]

  # Build full grid and fill NAs with None
  all_zips  <- sort(unique(zde$zip))
  all_dates <- as.IDate(seq(as.Date("2004-12-31"), as.Date("2022-12-31"), by = "day"))
  grid <- CJ(zip = as.character(all_zips), date = all_dates)

  effect_calendar <- zde[grid, on = .(zip, date)]
  effect_calendar[is.na(effect), effect := "None"]
  effect_calendar[, effect := factor(effect, levels = c("None","Cleanup","Threat","Impact"))]

  effect_calendar[]
}

# -------------------------------
# 2) Utility: run two_part_gcomp_bootstrap and return phase diffs vs None
# -------------------------------
run_by_threshold <- function(thr_knots,
                             Swaths_sf_full,
                             ER_visits_clean,
                             averaging_set = "exposed_today",
                             n_boot = 200,
                             cluster = "zip",
                             # -- Population controls (set these TRUE + pass zip_monthly if wanted)
                             pop_dt = NULL,
                             pop_by = "year_month",
                             use_poisson_offset = FALSE,
                             gcomp_weighting = "none") {

  ec <- build_effect_calendar_for_threshold(Swaths_sf_full, thr_knots)

  dat_model <- merge(
    ec[, .(zip, date, effect)],
    ER_visits_clean[, .(zip, date, er_visits)],
    by = c("zip","date"),
    all.x = TRUE
  )
  dat_model[is.na(er_visits), er_visits := 0L]

  res <- two_part_gcomp_bootstrap(
    df                 = dat_model,
    pop_dt             = pop_dt,              # pass zip_monthly if using population
    pop_by             = pop_by,              # "year_month" recommended
    use_poisson_offset = use_poisson_offset,  # TRUE for rate modeling
    gcomp_weighting    = gcomp_weighting,     # "population" to weight g-comp
    averaging_set      = averaging_set,       # "exposed_today" (your latest), or "ever_exposed"/"all"
    n_boot             = n_boot,
    cluster            = cluster,
    seed               = 123,
    verbose            = TRUE
  )

  out <- as.data.table(res$point_estimates$diff_vs_none)
  out[, threshold := paste0(thr_knots, " kt")]
  out[]
}

# -------------------------------
# 3) Run for 34, 55, and 64 kt and assemble tidy table
# -------------------------------
res34 <- run_by_threshold(34L, Swaths.sf, ER_visits_clean,
                          averaging_set = "ever_exposed",
                          n_boot = 200, cluster = "zip"
                          , pop_dt = zip_monthly, use_poisson_offset = TRUE, gcomp_weighting = "population"
)

res55 <- run_by_threshold(55L, Swaths.sf, ER_visits_clean,
                          averaging_set = "ever_exposed",
                          n_boot = 200, cluster = "zip"
                          , pop_dt = zip_monthly, use_poisson_offset = TRUE, gcomp_weighting = "population"
)

res64 <- run_by_threshold(64L, Swaths.sf, ER_visits_clean,
                          averaging_set = "ever_exposed",
                          n_boot = 200, cluster = "zip"
                          , pop_dt = zip_monthly, use_poisson_offset = TRUE, gcomp_weighting = "population"
)

effects_all <- rbindlist(list(res34, res55, res64), use.names = TRUE)

# Enforce desired plotting order Threat -> Impact -> Cleanup
effects_all[, effect := factor(effect, levels = c("Threat","Impact","Cleanup"))]
effects_all[, threshold := factor(threshold, levels = c("34 kt","55 kt","64 kt"))]

# -------------------------------
# 4) Make the 3-panel Figure 1
# -------------------------------
fig1 <- ggplot(effects_all, aes(x = effect, y = estimate)) +
  geom_hline(yintercept = 0, color = "grey50") +
  geom_pointrange(
    aes(ymin = ci_lo, ymax = ci_hi),
    color = "#2C3E50", linewidth = 0.6, fatten = 2
  ) +
  # ---- Change here: remove scales = "free_y" ----
  facet_wrap(~ threshold, nrow = 1) +
  labs(
    x = NULL,
    y = "Difference in expected ED visits vs None",
#    title = "Figure 1. Differences in expected ED visits vs None by storm phase and wind threshold",
#    subtitle = "Threat → Impact → Cleanup; estimand = exposed-today ZIP–days; n_boot = 200; clustered by #ZIP",
    caption = "Points: bootstrap means; bars: 95% percentile CIs"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title.position = "plot",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 10)
  )

dir.create("figs", showWarnings = FALSE)
ggsave(filename = "figs/Fig1_phase_effects_by_intensity.png",
       plot = fig1, width = 7, height = 5, dpi = 300)

effects_all[]
```

    effect   estimate      ci_lo      ci_hi threshold
    <fctr>      <num>      <num>      <num>    <fctr>
1: Cleanup  0.3435758  0.2811164  0.4097857     34 kt
2:  Threat  0.3187634  0.2656479  0.3849194     34 kt
3:  Impact -1.4517949 -1.5833675 -1.2872492     34 kt
4: Cleanup  0.6055326  0.4408778  0.7572212     55 kt
5:  Threat  0.2774763  0.1351130  0.3888371     55 kt
6:  Impact -2.7787372 -3.0691163 -2.5306009     55 kt
7: Cleanup  1.0085103  0.8131971  1.2175384     64 kt
8:  Threat  0.4973140  0.3418792  0.6858975     64 kt
9:  Impact -3.7167891 -4.0682611 -3.3906397     64 kt


